\pkg{partykit}: A Toolkit for Recursive Partytioning

### 2010-04-22 pkg/inst/doc/introduction.Rnw
playing with trees
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@164 255f4249-2263-4417-b1ea-8498779e27a7

<<mytree-3, echo = TRUE>>=
mytree <- function(formula, data, weights = NULL) {

@@

<<mytree-4, echo = TRUE>>=
aqt <- mytree(Ozone ~ Solar.R + Wind + Temp, data = airquality)
aqt
@

\begin{figure}
\begin{center}
<<mytree-5, echo = TRUE, fig = TRUE, width = 10, height = 6>>=
plot(aqt)
@
\caption{Tree}
\end{center}
\end{figure}

<<mytree-6, echo = TRUE>>=
predict(aqt, newdata = airquality[1:10,])
@

<<mytree-7, echo = TRUE>>=
aqt4 <- aqt[4]
aqt4
@

\begin{figure}
\begin{center}
<<mytree-8, echo = TRUE, fig = TRUE, width = 10, height = 6>>=
plot(aqt4)
@
\caption{Subtree}
\end{center}
\end{figure}

<<mytree-9, echo = TRUE>>=
info_split(split_node(node_party(aqt4)))
@

>>>>

A very rough sketch of formula-based user-interface needs
to set-up the data and call \code{growtree()}:
<<mytree-3, echo = TRUE>>=
mytree <- function(formula, data, weights = NULL) {

@@

We now can fit this tree, for example to the airquality data; the
\code{print()} method provides us with a first overview on the 
resulting model
<<mytree-4, echo = TRUE>>=
aqt <- mytree(Ozone ~ Solar.R + Wind + Temp, data = airquality)
aqt
@

\begin{figure}
\begin{center}
<<mytree-5, echo = TRUE, fig = TRUE, width = 10, height = 6>>=
plot(aqt)
@
\caption{Tree. \label{plottree}}
\end{center}
\end{figure}

We depict the model graphically using \code{plot()} (see Figure~\ref{plottree})
and compute predictions using
<<mytree-6, echo = TRUE>>=
predict(aqt, newdata = airquality[1:10,])
@

An interesting feature is the ability to extract subsets of trees:
<<mytree-7, echo = TRUE>>=
aqt4 <- aqt[4]
aqt4
@
which again are objects inheriting from \code{party} and thus
can be plotted easily (see Figure~\ref{subtree}).

\begin{figure}
\begin{center}
<<mytree-8, echo = TRUE, fig = TRUE, width = 10, height = 6>>=
plot(aqt4)
@
\caption{Subtree. \label{subtree}}
\end{center}
\end{figure}

We also might be interested in extracting the $p$-values in the
inner nodes in a nicely formatted way:
<<mytree-10, echo = TRUE>>=
fun <- function(x) 
    format.pval(info_split(split_node(x))$pvalue, 
                digits = 3, eps = 0.001)
nid <- nodeids(aqt)
iid <- nid[!(nid %in% nodeids(aqt, terminal = TRUE))]
unlist(nodeapply(aqt, ids = iid, FUN = fun))
@
